---
title: "p8105_hw2_kl3181"
author: Kelley Lou
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1
Read in Mr. Trashweel data set and clean.
```{r}
trashwheel_df = read_xlsx(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(
      sports_balls = round(sports_balls),
      sports_balls = as.integer(sports_balls)
  )
```

Read in precipitation data for 2018 and 2017. 
```{r}
precip_2018 = read_excel(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2018 Precipitation", 
      skip = 1
      ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = read_excel(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2017 Precipitation",
      skip = 1
      ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation dataframes. This can be done with a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge with precipitation dataset. 
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

precip_df = 
  left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr.Trashweel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashweel collects that trash, and then stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. There are variables including `r colnames(trashwheel_df)` and has data from `r min(pull(trashwheel_df, year))` to `r max(pull(trashwheel_df, year))`. The median number of sports balls in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`. The total precipitation in 2017 was `r precip_df %>% filter(year == 2017) %>% pull(total) %>% sum()` inches. In 2018, it was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2
#####Read in NYC transit dataset and clean.
```{r}
transit_df =  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
    janitor::clean_names() %>%
    select(line, station_name, station_latitude, station_longitude, route1:route11, entry, entrance_type, vending, ada) %>% 
    mutate(
      entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"
    ), 
)
```
This dataset focuses on NYC transit data, with information related to each entrance and exit for each subway station in NYC. After cleaning the dataset, the dataset contains variables `r colnames(transit_df)`. The clean_names function was utilized in the janitor package. The entry variable was changed from character to logical. Additionally, there are a total of `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. The data is not tidy because the route number and route name are not distinct variables.

##### Data Related Questions
```{r}
distinct_stations = distinct(transit_df, line, station_name, .keep_all = TRUE)
ada_compliant = nrow(filter(distinct_stations, ada == TRUE))
no_vending = transit_df %>% 
  filter(vending == "NO") 
prop = round(nrow(filter(no_vending, entry == TRUE)) / nrow(no_vending), 3)
```

There are `r nrow(distinct_stations)` distinct stations, `r ada_compliant` stations are ADA compliant and `r prop*100` percent of entrances/exits without vending allow entrance.

##### Reformat data
```{r}
transit_reformat = transit_df %>% 
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11, 
    names_to = "route_num", 
    values_to = "route_name") %>% 
  drop_na(route_name)

A_train = transit_reformat %>% 
  filter(route_name == "A") %>% 
  distinct(line, station_name, .keep_all = TRUE)

ADA_A = A_train %>% 
  filter(ada == TRUE)
```
There are `r nrow(A_train)` distinct stations serving the A train. Among these stations, `r nrow(ADA_A)` are ADA compliant.

