---
title: "p8105_hw2_kl3181"
author: Kelley Lou
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1
Read in Mr. Trashweel data set.
```{r}
trashwheel_df = read_xlsx(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "Mr. Trash Wheel",
      range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(
      sports_balls = round(sports_balls),
      sports_balls = as.integer(sports_balls)
  )
```

Read in precipitation data for 2018 and 2017. 
```{r}
precip_2018 = read_excel(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2018 Precipitation", 
      skip = 1
      ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = read_excel(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
      sheet = "2017 Precipitation",
      skip = 1
      ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation dataframes. This can be done with a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge with precipitation dataset. 
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

precip_df = 
  left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr.Trashweel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashweel collects that trash, and then stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. There are variables including `r colnames(trashwheel_df)` and has data from `r min(pull(trashwheel_df, year))` to `r max(pull(trashwheel_df, year))`. The median number of sports balls in 2017 was `r median(pull(filter(trashwheel_df, year == 2017), sports_balls))`. The total precipitation in 2017 was `r sum(pull(filter(precip_df, year == 2017), total))` inches. In 2018, it was `r sum(pull(filter(precip_df, year == 2018), total))` inches.


